---
title: 'Dataquest Guided Project: Finding the Best Markets to Advertise In'
author: "Cindy Zhang"
date: "11/30/2020"
output: 
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Introduction

This is my solution to Dataquest's Guided Project from the second Probability and Statistics course, which practices analyzing distributions of data from a fictional e-learning company.

More details such as the RMD and csv files can be found in the [repository in GitHub](https://github.com/chundychang/DQ-Rpracticeproject/tree/master/probability_statistics). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r pkg, include=FALSE}
library(knitr)
library(pander)
library(flextable)
library(magrittr)
library(dplyr)
library(purrr)
library(readr)
library(tidyverse)
library(tidyr)
library(tibble)
library(rlang)
library(DT)
library(stringr)
library(ggplot2)
options(stringsAsFactors = FALSE)
```

# Findings

## Step 1: Understanding the Data

```{r data, results='hide'}
coders <- data.frame(read_csv("coders.csv"))
dim(coders)
colnames(coders)
for (i in colnames(coders)) {
  print(class(coders[[i]]))
}
```

## Step 2: Checking for Sample Representativity

* Where are these new coders located.
* What locations have the greatest densities of new coders.
* How much money they're willing to spend on learning.

Clarify whether the data set has right categories of people for purpose. 

```{r jobrole}
coders %>%
  group_by(JobRoleInterest) %>%
  summarize(freq=n()*100/nrow(coders)) %>%
  arrange(desc(freq))
```

A lot interested in web development, some in mobile development, some in other parts; many interested in multiple subjects. How many interested in one versus many?

```{r split}
# split each string in the 'JobRoleInterest' column
split <- coders %>%
  select(JobRoleInterest) %>%
  tidyr::drop_na() %>%
  rowwise %>%
  mutate(opts = length(stringr::str_split(JobRoleInterest, ",")[[1]]))

# Frequency table for the var describing the number of options
n_of_options <- split %>%
  ungroup() %>%
  group_by(opts) %>%
  summarize(freq=n()*100/nrow(split))

n_of_options
```

Only 31.65% of participants have settled on one programming niche they'd like to work in; vast majority of students are mix. 
Focus of the course is web and mobile development --> find out how many respondents chose at least one of two options

```{r pop}
# Frequency table
web_or_mobile <- stringr::str_detect(coders$JobRoleInterest, "Web Developer|Mobile Developer")
freq_table <- table(web_or_mobile)
freq_table <- freq_table * 100 / sum(freq_table)
freq_table

#Graph for frequency table
df <- tibble::tibble(x=c("Other Subject", "Web or Mobile Development"),
                     y=freq_table)
ggplot(data=df, aes(x=x, y=y, fill=x)) + 
  geom_histogram(stat="identity")
```
Most survey repsondents (86%) are interested; suggests sample is representative of population; 

## Step 3: New Coders - Locations and Densities

```{r location}
freq_dist_location <- coders %>%
  tidyr::drop_na(JobRoleInterest)

freq_dist_location <- freq_dist_location %>%
  group_by(CountryLive) %>%
  summarize(Freq=n(), Percentage = n()*100/nrow(freq_dist_location)) %>%
  arrange(desc(Percentage))
```

44.69% of potential customers are in US; India has second customer density, but is only 7.55%, UK is 4.5%, Canada is 3.71%

## Step 4: Spending Money for Learning

```{r money}
# Replace 0s with 1s to avoid division by 0
mean_money <- coders %>%
  mutate(MonthsProgramming = replace(MonthsProgramming,  MonthsProgramming == 0, 1) )
# New column for the amount of money each student spends each month
mean_money <- mean_money %>%
  mutate(money_per_month = MoneyForLearning/MonthsProgramming)

mean_money %>%
  summarize(na_count=sum(is.na(money_per_month))) %>%
  pull(na_count)

#Keep only nonNA rows
mean_money <- mean_money %>%
  tidyr::drop_na(money_per_month)

#Remove NAs in CountryLive
mean_money <- mean_money %>%
  tidyr::drop_na(CountryLive)

#Frequency table to check if we still have enough data
mean_money %>%
  group_by(CountryLive) %>%
  summarize(freq=n()) %>%
  arrange(desc(freq)) %>%
  head()

```
```{r mean}
countries_mean <- mean_money %>%
  filter(CountryLive == 'United States of America' | CountryLive == 'India' | CountryLive == 'United Kingdom' | CountryLive == 'Canada') %>%
  group_by(CountryLive) %>%
  summarize(mean=mean(money_per_month)) %>%
  arrange(desc(mean))

countries_mean
```

Maybe we don't have enough representative data for the UK and Canada?

## Step 5: Dealing with Extreme Outliers

```{r}
#Isolate only the countries of interest, add index column
only_4 <- mean_money %>%
  filter(CountryLive == 'United States of America' | CountryLive == 'India' | CountryLive == 'United Kingdom'|CountryLive == 'Canada') %>%
  mutate(index = row_number())

#Box plots to visualize distributions
ggplot( data = only_4, aes(x = CountryLive, y = money_per_month)) +
  geom_boxplot() +
  ggtitle("Money Spent Per Month Per Country\n(Distributions)") +
  xlab("Country") +
  ylab("Money per month (US dollars)") +
  theme_bw()

```
we will need to remove those that spent exorbitant amounts of money (over $20,000)

```{r}
mean_money <- mean_money %>%
  filter(money_per_month < 20000)

# Mean sum of money spent by students each month
countries_mean <- mean_money %>%
  filter(CountryLive=='United States of America' | CountryLive == 'India' | CountryLive == 'United Kingdom'|CountryLive == 'Canada') %>%
  group_by(CountryLive) %>%
  summarize(mean=mean(money_per_month)) %>%
  arrange(desc(mean))

countries_mean
```

```{r}
# Isolate only the countries of interest
only_4  <-  mean_money %>% 
  filter(CountryLive == 'United States of America' | CountryLive == 'India' | CountryLive == 'United Kingdom'|CountryLive == 'Canada') %>%
  mutate(index = row_number())

# Box plots to visualize distributions
ggplot( data = only_4, aes(x = CountryLive, y = money_per_month)) +
  geom_boxplot() +
  ggtitle("Money Spent Per Month Per Country\n(Distributions)") +
  xlab("Country") +
  ylab("Money per month (US dollars)") +
  theme_bw()

```
some extreme outliers for India (values over $2,500 a month); did these people attend several conferences, which tend to be expensive?

```{r}
#Inspect India outliers

india_outliers <- only_4 %>%
  filter(CountryLive == 'India' & 
           money_per_month >= 2500)

india_outliers
```

None attended bootcamp, best to remove outliers

```{r}
only_4 <- only_4 %>%
  filter(!(index %in% india_outliers$index))
```

Some extreme outliers for US

```{r}
#examine US extreme outliers
us_outliers = only_4 %>%
    filter(CountryLive == 'United States of America' & 
           money_per_month >= 6000)

us_outliers

only_4 <- only_4 %>%
  filter(!(index %in% us_outliers$index))
```

Out of these 11 extreme outliers, six people attended bootcamps, which justify the large sums of money spent on learning. For the other five, it's hard to figure out from the data where they could have spent that much money on learning. Consequently, we'll remove those rows where participants reported thay they spend \$6,000 each month, but they have never attended a bootcamp.

Also, the data shows that eight respondents had been programming for no more than three months when they completed the survey. They most likely paid a large sum of money for a bootcamp that was going to last for several months, so the amount of money spent per month is unrealistic and should be significantly lower (because they probably didn't spend anything for the next couple of months after the survey). As a consequence, we'll remove every these eight outliers.

```{r}
#Remove respondents who didn't attend bootcamp
no_bootcamp = only_4 %>%
    filter(CountryLive == 'United States of America' & 
           money_per_month >= 6000 &
             AttendedBootcamp == 0)

only_4 <- only_4 %>%
  filter(!(index %in% no_bootcamp$index))

# Remove the respondents that had been programming for less than 3 months
less_than_3_months = only_4 %>%
  filter(CountryLive == 'United States of America' &
           money_per_month >= 6000 &
           MonthsProgramming <= 3)

only_4  <-  only_4 %>% 
  filter(!(index %in% less_than_3_months$index))
```

Looking again at the last box plot above, we can also see an extreme outlier for Canada — a person who spends roughly \$5,000 per month. Let's examine this person in more depth.

```{r}
#Examine the extreme outliers for Canada
canada_outliers = only_4 %>%
  filter(CountryLive == 'Canada' &
           money_per_month >= 4500 &
           MonthsProgramming <= 3)

canada_outliers
```
Here, the situation is similar to some of the US respondents — this participant had been programming for no more than two months when he completed the survey. He seems to have paid a large sum of money in the beginning to enroll in a bootcamp, and then he probably didn't spend anything for the next couple of months after the survey. We'll take the same approach here as for the US and remove this outlier.

```{r}
#Remove Canada extreme outlier
only_4  <-  only_4 %>% 
  filter(!(index %in% canada_outliers$index))
```

```{r}
#Mean sum of money spent by students each month, no outliers
countries_mean = only_4 %>%
  group_by(CountryLive) %>%
  summarize(mean=mean(money_per_month)) %>%
  arrange(desc(mean))

countries_mean
```
```{r}
#Box plots to visualize distributions
ggplot( data = only_4, aes(x = CountryLive, y = money_per_month)) +
  geom_boxplot() +
  ggtitle("Money Spent Per Month Per Country\n(Distributions)") +
  xlab("Country") +
  ylab("Money per month (US dollars)") +
  theme_bw()
```

## Step 6: Choosing the Two Best Markets

Obviously, one country we should advertise in is the US. Lots of new coders live there and they are willing to pay a good amount of money each month (roughly \$143).

We sell subscriptions at a price of \$59 per month, and Canada seems to be the best second choice because people there are willing to pay roughly \$93 per month, compared to India (\$66) and the United Kingdom (\$45).

The data suggests strongly that we shouldn't advertise in the UK, but let's take a second look at India before deciding to choose Canada as our second best choice:

* $59 doesn't seem like an expensive sum for people in India since they spend on average \$66 each month.
* We have almost twice as more potential customers in India than we have in Canada:

```{r}
# Frequency table for the 'CountryLive' column
only_4 %>%
  group_by(CountryLive) %>%
  summarize(freq=n()*100/nrow(only_4)) %>%
  arrange(desc(freq)) %>%
  head()
```
```{r}
#Frequency table check if enough data
only_4 %>%
  group_by(CountryLive) %>%
  summarize(freq=n()) %>%
  arrange(desc(freq)) %>%
  head()
```

So it's not crystal clear what to choose between Canada and India. Although it seems more tempting to choose Canada, there are good chances that India might actually be a better choice because of the large number of potential customers.

At this point, it seems that we have several options:

1. Advertise in the US, India, and Canada by splitting the advertisement budget in various combinations:
    - 60% for the US, 25% for India, 15% for Canada.
    - 50% for the US, 30% for India, 20% for Canada; etc.

2. Advertise only in the US and India, or the US and Canada. Again, it makes sense to split the advertisement budget unequally. For instance:
    - 70% for the US, and 30% for India.
    - 65% for the US, and 35% for Canada; etc.

3. Advertise only in the US.

At this point, it's probably best to send our analysis to the marketing team and let them use their domain knowledge to decide. They might want to do some extra surveys in India and Canada and then get back to us for analyzing the new survey data.

# Conclusion

In this project, we analyzed survey data from new coders to find the best two markets to advertise in. The only solid conclusion we reached is that the US would be a good market to advertise in.

For the second best market, it wasn't clear-cut what to choose between India and Canada. We decided to send the results to the marketing team so they can use their domain knowledge to take the best decision.