---
title: 'Dataquest Guided Project: Analyzing Movie Ratings'
author: "Cindy Zhang"
date: "2/6/2021"
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---
# Introduction

This is my solution to Dataquest's Guided Project from the Web Scraping in R course, which practices web scraping in R.

More details such as the RMD and csv files can be found in the [repository in GitHub](https://github.com/chundychang/DQ-Rpracticeproject/tree/master/api_web_scrape). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r pkg, include=FALSE}
library(knitr)
library(pander)
library(httr)
library(flextable)
library(magrittr)
library(dplyr)
library(purrr)
library(readr)
library(tidyverse)
library(tidyr)
library(tibble)
library(rlang)
library(DT)
library(stringr)
library(ggplot2)
library(rvest)
options(stringsAsFactors = FALSE)
```

# Findings

## Loading the Web Page

```{r}
wp_content <- read_html("http://dataquestio.github.io/web-scraping-pages/IMDb-DQgp.html")
```

## Extracting Elements

```{r}
titles <- wp_content %>%
  html_nodes("h3 a") %>%
  html_text()
years <- wp_content %>%
  html_nodes(".lister-item-year") %>%
  html_text()

#convert years from character to numeric
years <- readr::parse_number(years)
```

## Extracting Movies' Features

```{r}
runtimes <- wp_content %>%
  html_nodes(".runtime") %>%
  html_text()
genres <- wp_content %>%
  html_nodes(".genre") %>%
  html_text()

# Conversion/cleaning

runtimes <- readr::parse_number(runtimes)
genres <- str_sub(genres, 2, -1)
genres <- stringr::str_trim(genres)
```

## Extracting Movies' Ratings

```{r}
user_ratings <- wp_content %>%
  html_nodes("div .ratings-imdb-rating") %>%
  html_text
metascores <- wp_content %>%
  html_nodes(".metascore") %>%
  html_text()

# Cleaning
metascores <- readr::parse_number(metascores)
user_ratings <- str_replace_all(user_ratings, "\n", "")
user_ratings<- str_trim(user_ratings)
user_ratings <- readr::parse_number(user_ratings)
```

## Extracting Movies' Votes

```{r}
votes <- wp_content %>%
  html_nodes(".sort-num_votes-visible") %>%
  html_text()

# Cleaning
votes <- str_replace_all(votes, "\n                Votes:\n                ", "")
votes <- str_replace_all(votes, "\n", "")
votes <- str_trim(votes)
votes <- readr::parse_number(votes)
```

## Dealing with Missing Values

```{r}
append_vector <- function(vector, inserted_indices, values){

  ## Creating the current indices of the vector
  vector_current_indices <- 1:length(vector)

  ## Adding small amount of values (between 0 and 0.9) to the `inserted_indices`
  new_inserted_indices <- inserted_indices + seq(0, 0.9, length.out = length(inserted_indices))

  ## Appending the `new_inserted_indices` to the current vector indices
  indices <- c(vector_current_indices, new_inserted_indices)

  ## Ordering the indices
  ordered_indices <- order(indices)

  ## Appending the new value to the existing vector
  new_vector <- c(vector, values)

  ## Ordering the new vector wrt the ordered indices
  new_vector[ordered_indices]
}
metascores <- append_vector(metascores, c(1,1,1,13,24), NA)
titles <- titles[-17]
years <- years[-17]
runtimes <- runtimes[-17]
genres <- genres[-17]
metascores <- metascores[-17]
```

## Visualization

```{r}
ratings_df <- tibble(titles, years, runtimes, genres, user_ratings, metascores, votes)
# Round user ratings to integers
ratings_df <- ratings_df %>%
  mutate(user_ratings=floor(user_ratings))
# Visualize
ggplot(data=ratings_df,
       aes(x=user_ratings,y=votes,group=user_ratings))+geom_boxplot()
```
The lowest- and highest-rated films have the lowest and highest votes, respectively. Linnear correlation between ratings and votes.
