---
title: 'Dataquest Guided Project: Analyzing Movie Ratings'
author: "Cindy Zhang"
date: "2/6/2021"
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---
# Introduction

This is my solution to Dataquest's Guided Project from the Web Scraping in R course, which practices web scraping in R.

More details such as the RMD and csv files can be found in the [repository in GitHub](https://github.com/chundychang/DQ-Rpracticeproject/tree/master/api_web_scrape). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r pkg, include=FALSE}
library(knitr)
library(pander)
library(httr)
library(flextable)
library(magrittr)
library(dplyr)
library(purrr)
library(readr)
library(tidyverse)
library(tidyr)
library(tibble)
library(rlang)
library(DT)
library(stringr)
library(ggplot2)
library(rvest)
options(stringsAsFactors = FALSE)
```

# Findings

## Loading the Web Page

```{r}
wp_content <- read_html("http://dataquestio.github.io/web-scraping-pages/IMDb-DQgp.html")
```

## Extracting Elements

```{r}
titles <- wp_content %>%
  html_nodes("h3 a") %>%
  html_text()
years <- wp_content %>%
  html_nodes("span ")
```

## Extracting Movies' Features

## Extracting Movies' Ratings

## Extracting Movies' Votes

## Dealing with Missing Values

## Visualization
