---
title: 'Dataquest Guided Project: New York Solar Resource Data'
author: "Cindy Zhang"
date: "1/22/2021"
output: 
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Introduction

This is my solution to Dataquest's Guided Project from the APIs in R course, which practices querying APIs in R.

More details such as the RMD and csv files can be found in the [repository in GitHub](https://github.com/chundychang/DQ-Rpracticeproject/tree/master/api_web_scrape). 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r pkg, include=FALSE}
library(knitr)
library(pander)
library(httr)
library(flextable)
library(magrittr)
library(dplyr)
library(purrr)
library(readr)
library(tidyverse)
library(tidyr)
library(tibble)
library(rlang)
library(DT)
library(stringr)
library(ggplot2)
options(stringsAsFactors = FALSE)
```

# Findings

## API set up

This guided project involves querying the Data Gov API, which requires an access key. Enter your own data key; if you don't have one, you can sign up for one [here] (https://api.data.gov/signup/). 

```{r, echo=FALSE}
# storing the values

# make sure you hide your API key when publishing with echo=FALSE

key <- #enter your API key here.
```

First, I stored my endpoint of interest and parameters in a variable and a list, respectively. Then I wrote a function to query the Solar data API and executed the function, storing the results.

```{r}
solar_url <- "/api/solar/solar_resource/v1"
parameters_list <- (list(format = "json", api_key = key, lat = "41", lon = "-75"))

#api get function
gov_api_json <- function(endpoint, queries = list()) {

  # Preparing the URL 
  url <- modify_url("https://developer.nrel.gov", path = endpoint)

  # API requests
  response <- GET(url, query = queries)

  # Tracking errors
  if ( http_error(response) ){
    print(status_code(response))
    stop("Something went wrong.", call. = FALSE)
  }

  if (http_type(response) != "application/json") {
    stop("API did not return json", call. = FALSE)
  }

  # Extracting content
  json_text <- content(response, "text")

  # Store content as list
  json_lists <- jsonlite::fromJSON(json_text)
  
  # Return list
  json_lists
}

# Execute function

complex_list <- gov_api_json(solar_url, queries = parameters_list)
str(complex_list)

```

## Create dataframe from complex list

Since `complex_list` is a bunch of nested lists, I converted it into a work-able dataframe. I had to manually re-add the month columns. 

```{r}
# simplify list- avg_dni
u_complex_list <- unlist(complex_list$outputs)

# restructure the new list into a matrix
u_df <- as.data.frame(matrix(u_complex_list, nrow=13))

# remove first row 
u_df <- u_df[-1,]

# rename columns
u_df <- u_df %>%
  rename(
    avg_dni = V1,
    avg_ghi = V2,
    avg_lat_tilt = V3
  )

# add month column
u_df <- u_df %>%
  mutate(month = month.abb) %>%
  select(month, everything())

```
