---
title: 'Dataquest Guided Project: Predicting Condominium Sale Prices'
author: "Cindy Zhang"
date: "8/9/2021"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Introduction

This is my solution to Dataquest's Guided Project from the first Predictive Modeling and Machine Learning in R course, which investigates how well condominium sizes predict sale prices in New York City as a whole and each of the five boroughs.

More details such as the RMD and csv files can be found in the [repository in GitHub](https://github.com/chundychang/DQ-Rpracticeproject/tree/master/probability_statistics).


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r pkg, include=FALSE}
library(knitr)
library(pander)
library(httr)
library(flextable)
library(magrittr)
library(dplyr)
library(purrr)
library(readr)
library(tidyverse)
library(tidyr)
library(tibble)
library(rlang)
library(DT)
library(stringr)
library(ggplot2)
library(rvest)
library(broom)
options(stringsAsFactors = FALSE)
```

# Findings

## Understanding the Data

```{r}
#Load data
NYC_property_sales <- suppressMessages(read_csv("NYC_property_sales.csv"))

#building classifications
sort(table(NYC_property_sales$building_class_at_present))

```

## Explore Bivariate Relationships with Scatterplots

This project focuses on condominiums with elevators, classified as "R4." So the first step is subsetting the data to include only properties classified as condominiums with elevators at time of sale:

```{r}
NYC_condos <- NYC_property_sales %>%
  filter(building_class_at_time_of_sale=="R4")
```

Next, I generate a scatterplot to visualize the relationship between `gross_square_feet` and `sale_price`.

```{r}
ggplot(data=NYC_condos, 
       aes(x=gross_square_feet, y=sale_price))+
  geom_point(aes(color=borough), alpha=0.3)+
  scale_y_continuous(labels=scales::comma, limits=c(0,75000000))+
  xlim(0,10000)+
  geom_smooth(method="lm", se=FALSE)+
  labs(title="Condominium Sale Price Explained by Gross Square Feet",
       x="Gross Square Feet (sqft)",
       y="Sale Price($)")
```
The scatterplot generated shows that most of the data is clustered towards the bottom left of the graph. I adjust the x and y limits to zoom in and space out the points.

```{r}
ggplot(data=NYC_condos, 
       aes(x=gross_square_feet, y=sale_price))+
  geom_point(aes(color=borough), alpha=0.3)+
  scale_y_continuous(labels=scales::comma, limits=c(0,20000000))+
  xlim(0,5000)+
  geom_smooth(method="lm", se=FALSE)+
  labs(title="Condominium Sale Price Explained by Gross Square Feet",
       x="Gross Square Feet (sqft)",
       y="Sale Price($)")
```

I then use `facet_wrap()` to split the graph into boroughs so I can view the spread of data for each borough individually.

```{r}
ggplot(data=NYC_condos, 
       aes(x=gross_square_feet, y=sale_price))+
  geom_point(alpha=0.3)+
  facet_wrap(~borough, scales="free", ncol=2)+
  scale_y_continuous(labels=scales::comma)+
  geom_smooth(method="lm", se=FALSE)+
  labs(title="Condominium Sale Price Explained by Gross Square Feet per Borough",
       x="Gross Square Feet (sqft)",
       y="Sale Price($)")
```

## Outliers and Data Integrity Issues

As shown by the faceted graphs, there are several outliers that will need to be removed, specifically multi-unit sales that misrepresent sale price. I remove several below:

```{r}
#save original subsetted data
NYC_condos_original <- NYC_condos

#remove 165 East 66
NYC_condos <- NYC_condos %>%
  filter(address != "165 East 66th St, Resi")

# Investigating multi unit sales
multi_unit_sales <- NYC_condos %>% 
  group_by(sale_price, sale_date) %>% 
  filter(n() >= 3) %>% 
  arrange(desc(sale_price))

# Removing multi unit sales
NYC_condos <- NYC_condos %>%
  group_by(sale_price, sale_date) %>%
  filter(n() <= 2) %>%
  ungroup()
```

## Linear Regression Model for Boroughs in New York City Combined

```{r}
# Generate linear model of sale price explained by gross square feet
NYC_condos_lm <- lm(sale_price~gross_square_feet, data=NYC_condos)

# Generate linear model of sale price epxlained by gross square feet with original data
NYC_condos_original_lm <- lm(sale_price~gross_square_feet,data=NYC_condos_original)

summary(NYC_condos_original_lm)
summary(NYC_condos_lm)

```
The linear models test the following hypotheses:
* H0: There is no significant relationship between sale price and gross square feet (the slope is zero)
* Ha: There is a significant relationship between sale price and gross square feet (the slope is far from zero)

For both models, the t-stat was high enough and the p-value low enough to determine a significant relationship between sale price and gross square feet for both models. Notably, the t-stat for the cleaned dataset `NYC_condos` was nearly double that of the original dataset `NYC_condos_original`. Both p-values are well below 0.05, making it unlikely that the relationship between sale price and gross square feet is due to random chance. 

```{r}
confint(NYC_condos_original_lm)[2,]
confint(NYC_condos_lm)[2,]
```
However, the two models have very different confidence intervals, likely an effect of the outliers present in the original model and not present in the cleaned model.

```{r}
sigma(NYC_condos_original_lm)
sigma(NYC_condos_lm)
```
Furthermore, the original model has more of a lack of fit compared to the cleaned model. The cleaned model also has a higher R-squared value than the original model. 

```{r}
ggplot(data=NYC_condos, 
       aes(x=gross_square_feet, y=sale_price))+
  geom_point(alpha=0.3)+
  facet_wrap(~borough, scales="free", ncol=2)+
  scale_y_continuous(labels=scales::comma)+
  geom_smooth(method="lm", se=FALSE)+
  labs(title="Condominium Sale Price Explained by Gross Square Feet per Borough",
       x="Gross Square Feet (sqft)",
       y="Sale Price($)")
```
It's now easier to see the spread of the data, especially for Brooklyn and Manhattan.

## Linear Regression Models - Coefficient Estimates

Next, I compare linear models of `sale_price` predicted by `gross_square_feet` for each borough individually.

```{r}
# Nest by borough
NYC_nested <- NYC_condos %>%
  group_by(borough) %>%
  nest()

# Fit linear models to each borough individually
NYC_coefficients <- NYC_condos %>%
  group_by(borough) %>%
  nest() %>%
  mutate(linear_model=map(.x=data,
                          .f=~lm(sale_price~gross_square_feet,
                          data=.)))

# Test
summary(NYC_coefficients$linear_model[[2]])

# Tidy with Confidence Intervals

NYC_tidy <- NYC_condos %>%
  group_by(borough) %>%
  nest() %>%
  mutate(linear_model=map(.x=data,
                          .f=~lm(sale_price~gross_square_feet,
                          data=.))) %>%
  mutate(tidy_coefficients=map(.x=linear_model,
                               .f=tidy,
                               conf.int=TRUE))

NYC_tidy$tidy_coefficients[[4]]

# Unnest to tidy dataframe

NYC_coefficients_tidy <- NYC_tidy %>%
  select(borough, tidy_coefficients) %>%
  unnest(cols=tidy_coefficients)
print(NYC_coefficients_tidy)

# Return slope estimate only
NYC_slope <- NYC_coefficients_tidy %>%
  filter(term=="gross_square_feet") %>%
  arrange(estimate)

print(NYC_slope)
```

Based on `NYC_slope`, the t-stats and p-values for all five boroughs indicate a significat relationship between sale price and gross square feet. In Staten Island, an increase in total square footage by one unit increases sale price by about 289 dollars, in contrast, the same increase increases sale price by 4,728 dollars.

## Linear Regression Models for each Borough - Regression Summary Statistics

```{r}
NYC_nested <- NYC_condos %>%
  group_by(borough) %>%
  nest() %>%
  mutate(linear_model=map(.x=data,
                          .f=~lm(sale_price~gross_square_feet,
                          data=.))) %>%
  mutate(tidy_coefficients=map(.x=linear_model,
                               .f=tidy,
                               conf.int=TRUE)) %>%
  mutate(tidy_summary_stats=map(.x=linear_model,
                                .f=glance))
NYC_summary_stats <- NYC_nested %>%
  select(borough, tidy_summary_stats) %>%
  unnest(cols=tidy_summary_stats)

print(NYC_summary_stats)
```

# Conclusion

In general, `gross_square_feet` is useful for explaining/estimating `sale_price` for condos in NYC. However, there is a wide range in slope estimates by borough. For example, an increase in square footage yields a greater increase in price in Manhattan than in Staten Island. Additionally, `gross_square_feet` is a better single predictor of `sale_price` in some boroughs than in others. For example, Manhattan has a higher adjusted R squared than the other boroughs, suggesting that total square footage is more effective at predicting sale price in Manhattan than the other boroughs. 