---
title: 'Dataquest Guided Project: NYC Schools Perceptions'
author: "Cindy Zhang"
date: "8/22/2020"
output: 
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Introduction

This is my solution to Dataquest's Guided Project from the first Data Cleaning in R course, which analyzes survey responses on quality perceptions of NYC schools. This analysis focuses on answering two questions:

1. Do student, teacher, and parent perceptions of NYC school quality appear to be related to demographic and academic success metrics?
2. Do students, teachers, and parents have similar perceptions of NYC school quality? 

More details such as the RMD and csv files can be found in the [repository in GitHub](https://github.com/chundychang/DQ-Rpracticeproject/tree/master/nyc%20schools). Details about the variables can be found [here](https://data.cityofnewyork.us/Education/2011-NYC-School-Survey/mnz3-dyi8).


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r pkg, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(purrr)
library(readr)
library(tidyverse)
library(tibble)
library(DT)
library(stringr)
library(lubridate)
library(ggplot2)
options(stringsAsFactors = FALSE)
```

# Findings

## Step 1: Loading the Dataframes

```{r load}
combined <- data.frame(read_csv("combined.csv"))
gened <- data.frame(read_tsv("masterfile11_gened_final.txt"))
d75 <- data.frame(read_tsv("masterfile11_d75_final.txt"))
```

## Step 2: Simplifying the Dataframes

dbn, high school, location, enrollment, borough, columns 14-32

```{r simplify, echo=FALSE}
gened_simp <- gened %>%
  select(dbn, bn, studentssurveyed, schoolname, highschool, 14:32) %>%
  filter(is.na(highschool))

d75_simp <- d75 %>%
  select(dbn, bn, studentssurveyed, schoolname, highschool, 14:32) %>%
  filter(is.na(highschool))
```

## Step 3: Creating a Single Dataframe for Analysis

```{r combine}
surveys <- bind_rows(gened, d75)
combined_surv <- combined %>%
  rename(dbn = DBN) %>%
  left_join(surveys, by = "dbn")
```

## Step 4: Looking for Interesting Correlations and Examining Relationships Using Scatter Plots

```{r cor}

```

## Step 5: Reshaping the Data Based on Differences in Student,Parent, and Teacher Perceptions
