---
title: 'Dataquest Guided Project: Designing and Creating a Database'
author: "Cindy Zhang"
date: "10/10/2020"
output: 
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Introduction

This is my solution to Dataquest's Guided Project from the Intermediate SQL in R course, which practices designing a database using SQLite. 

More details such as the RMD and database files can be found in the [repository in GitHub](https://github.com/chundychang/DQ-Rpracticeproject/tree/master/SQL%20projects).

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r pkg, include=FALSE}
library(knitr)
library(forcats)
library(RSQLite)
library(DBI)
library(pander)
library(flextable)
library(magrittr)
library(dplyr)
library(purrr)
library(readr)
library(tidyverse)
library(tidyr)
library(tibble)
library(rlang)
library(DT)
library(stringr)
library(ggplot2)
options(stringsAsFactors = FALSE)
```

# Findings

## Step 1: Exploring the Data

## Step 2: Importing Data into SQLite

## Step 3: looking for Normalization Opportunities

## Step 4: Planning a Normalized Schema

## Step 5: Creating Tables Without Foreign Key Relations

## Step 6: Adding the Team and Game Tables

## Step 7: Adding the Team Appearance Table

## Step 8: Adding the Person Appearance Table

## Step 9: Removing the Original Tables

# Conclusion