---
title: 'Dataquest Guided Project: Designing and Creating a Database'
author: "Cindy Zhang"
date: "10/10/2020"
output: 
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Introduction

This is my solution to Dataquest's Guided Project from the Intermediate SQL in R course, which practices designing a database using SQLite. 

More details such as the RMD and database files can be found in the [repository in GitHub](https://github.com/chundychang/DQ-Rpracticeproject/tree/master/SQL%20projects). Additional information can be found in (https://www.retrosheet.org/eventfile.htm)

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

```{r pkg, include=FALSE}
library(knitr)
library(forcats)
library(RSQLite)
library(sqldf)
library(data.table)
library(DBI)
library(pander)
library(flextable)
library(magrittr)
library(dplyr)
library(purrr)
library(readr)
library(tidyverse)
library(tidyr)
library(tibble)
library(rlang)
library(DT)
library(stringr)
library(ggplot2)
options(stringsAsFactors = FALSE)
```

# Findings

## Step 1: Exploring the Data

```{r data_load}
park_codes_df <- data.frame(read_csv("park_codes.csv"))
person_codes_df <- data.frame(read_csv("person_codes.csv"))
team_codes_df <- data.frame(read_csv("team_codes.csv"))
```

```{r game_log_load}
game_log_df <- read_csv("game_log.csv", 
                col_types = cols(.default = "c",
                                 v_league = "c", h_league = "c",
                                 `3b_umpire_id` = "c", `3b_umpire_name` = "c",
                                 `2b_umpire_id` = "c", `2b_umpire_name` = "c",
                                 `lf_umpire_id` = "c", `lf_umpire_name` = "c",
                                 `rf_umpire_id` = "c", `rf_umpire_name` = "c",
                                 completion = "c", winning_rbi_batter_id = "c",
                                 winning_rbi_batter_id_name = "c", protest = "c",
                                 v_first_catcher_interference = "c", 
                                 h_first_catcher_interference = "c"))
```

```{r EDA_1, echo=FALSE}
dim(game_log_df)
colnames(game_log_df)
for (i in colnames(game_log_df)) {
  print(class(game_log_df[[i]]))
}
```

What each defensive position number represents
1 = Pitcher
2 = Catcher
3 = First baseman
4 = second baseman
5 = third baseman
6 = shortstop
7 = left fielder
8 = center fielder
9 = right fielder

```{r EDA_2, echo=FALSE}
dim(park_codes_df)
colnames(park_codes_df)
for (i in colnames(park_codes_df)) {
  print(class(park_codes_df[[i]]))
}
```

```{r EDA_3, echo=FALSE}
dim(person_codes_df)
colnames(person_codes_df)
for (i in colnames(person_codes_df)) {
  print(class(person_codes_df[[i]]))
}
```

```{r EDA_4, echo=FALSE}
dim(team_codes_df)
colnames(team_codes_df)
for (i in colnames(team_codes_df)) {
  print(class(team_codes_df[[i]]))
}
```

## Step 2: Importing Data into SQLite

```{r helper}
conn <- dbConnect(SQLite(), dbname="mlb")
dbWriteTable(conn = conn, name = "game_log", value = game_log_df, header = TRUE, overwrite = TRUE)
dbWriteTable(conn = conn, name = "park_codes", value = park_codes_df, header = TRUE, overwrite = TRUE)
dbWriteTable(conn = conn, name = "person_codes", value = person_codes_df, header = TRUE, overwrite = TRUE)
dbWriteTable(conn = conn, name = "team_codes", value = team_codes_df, header = TRUE, overwrite = TRUE)
dbListTables(conn)
```

```{r game_id}
dbExecute(conn,
          "ALTER TABLE game_log
          ADD COLUMN game_id TEXT")
```
```{r game_id_1}
dbExecute(conn,
          "UPDATE game_log
          SET game_id = h_name || date || number_of_game;")
```
## Step 3: Creating Tables Without Foreign Key Relations

```{r person}
dbExecute(conn,
          "CREATE TABLE person (
          person_id TEXT PRIMARY KEY,
          first_name TEXT,
          last_name TEXT);")
dbExecute(conn,
          "INSERT INTO person
          SELECT id, first, last
          FROM person_codes")
dbExecute(conn,
          "SELECT *
          FROM person
          LIMIT 5")
```

```{r park}
dbExecute(conn, 
          "CREATE TABLE park (
          park_id TEXT PRIMARY KEY,
          name TEXT,
          nickname TEXT,
          city TEXT,
          state TEXT,
          notes TEXT);")
dbExecute(conn,
          "INSERT INTO park
          SELECT park_id, name, aka, city, state, notes
          FROM park_codes")
dbExecute(conn,
          "SELECT *
          FROM park
          LIMIT 5")
```

```{r league}
dbExecute(conn,
          "CREATE TABLE league (
          league_id TEXT PRIMARY KEY,
          name TEXT);")
dbExecute(conn,
          'INSERT INTO league
          VALUES
          ("AL", "American League"),
          ("AA", "American Association"),
          ("FL", "Federal League"),
          ("NL", "National League"),
          ("PL", "Players League"),
          ("UA", "Union Association");')
dbExecute(conn,
          "SELECT *
          FROM league")
```

League information was found here: http://www.sportsforecaster.com/mlb/abbreviations

```{r appearance_type}
appearance_type_df <- data.frame(read_csv("appearance_type.csv"))
dbWriteTable(conn = conn, name = "appearance_type", value = appearance_type_df, header = TRUE, overwrite = TRUE)
dbListTables(conn)
```

## Step 6: Adding the Team and Game Tables

```{r team}
dbExecute(conn,
          "CREATE TABLE team (
          team_id TEXT,
          league_id TEXT,
          city TEXT,
          nickname TEXT,
          franch_id TEXT,
          PRIMARY KEY (team_id, league_id)
          FOREIGN KEY (league_id) REFERENCES league(league_id));")
dbExecute(conn,
          "INSERT INTO team
          SELECT team_id, league, city, nickname, franch_id
          FROM team_codes")
dbExecute(conn,
          "SELECT *
          FROM team
          LIMIT 5")
```

```{r game}
dbExecute(conn,
          "CREATE TABLE game (
          game_id TEXT PRIMARY KEY, 
          date DATETIME,
          number_of_game INTEGER,
          park_id TEXT, 
          length_outs INTEGER,
          day BLOB, 
          completion TEXT,
          forfeit TEXT,
          protest TEXT,
          attendance INTEGER,
          length_minutes INTEGER,
          additional_info TEXT,
          acquisition_info TEXT,
          FOREIGN KEY (park_id) REFERENCES park(park_id))")
dbExecute(conn,
          "INSERT INTO game
          SELECT game_id, date, number_of_game, park_id, length_outs, day_of_week, completion, forefeit, protest, attendance, length_minutes, additional_info, acquisition_info
          FROM game_log")
dbExecute(conn,
          "SELECT *
          FROM game
          LIMIT 5")
```

## Step 7: Adding the Team Appearance Table

```{r team_appearance_create}
dbExecute(conn,
          "CREATE TABLE team_appearance (
          team_id TEXT,
          game_id TEXT,
          home BLOB,
          league_id TEXT,
          score INTEGER,
          line_score INTEGER,
          at_bats INTEGER,
          hits INTEGER,
          doubles INTEGER,
          triples INTEGER,
          homeruns INTEGER,
          rbi INTEGER,
          sacrifice_hits INTEGER,
          sacrifice_flies INTEGER,
          hit_by_pitch INTEGER,
          walks INTEGER,
          intentional_walks INTEGER,
          strikeouts INTEGER,
          stolen_bases INTEGER,
          caught_stealing INTEGER,
          grounded_into_double INTEGER,
          first_catcher_interference INTEGER,
          left_on_base INTEGER,
          pitchers_used INTEGER,
          individual_earned_runs INTEGER,
          team_earned_runs INTEGER,
          wild_pitches INTEGER,
          balks INTEGER,
          putouts INTEGER,
          assists INTEGER,
          errors INTEGER,
          passed_balls INTEGER,
          double_plays INTEGER,
          triple_plays INTEGER,
          PRIMARY KEY (team_id, game_id),
          FOREIGN KEY (team_id) REFERENCES team(team_id),
          FOREIGN KEY (game_id) REFERENCES game(game_id),
          FOREIGN KEY (league_id) REFERENCES league(league_id))")
```

```{r team_appearance_insert}
dbExecute(conn, 
          "INSERT INTO team_appearance
            SELECT h_name, game_id, 1 AS home, h_league, h_score, h_line_score, h_at_bats, h_hits, h_doubles, h_triples, h_homeruns, h_rbi, h_sacrifice_hits, h_sacrifice_flies, h_hit_by_pitch, h_walks, h_intentional_walks, h_strikeouts, h_stolen_bases, h_caught_stealing, h_grounded_into_double, h_first_catcher_interference, h_left_on_base, h_pitchers_used, h_individual_earned_runs, h_team_earned_runs, h_wild_pitches, h_balks, h_putouts, h_assists, h_errors, h_passed_balls, h_double_plays, h_triple_plays
            FROM game_log
          UNION
            SELECT v_name, game_id, 0 AS home, v_league, v_score, v_line_score, v_at_bats, v_hits, v_doubles, v_triples, v_homeruns, v_rbi, v_sacrifice_hits, v_sacrifice_flies, v_hit_by_pitch, v_walks, v_intentional_walks, v_strikeouts, v_stolen_bases, v_caught_stealing, v_grounded_into_double, v_first_catcher_interference, v_left_on_base, v_pitchers_used, v_individual_earned_runs, v_team_earned_runs, v_wild_pitches, v_balks, v_putouts, v_assists, v_errors, v_passed_balls, v_double_plays, v_triple_plays
          FROM game_log;")
```

## Step 8: Adding the Person Appearance Table

## Step 9: Removing the Original Tables

# Conclusion